---
title: "A Network Analysis of Global Preferential Trade Agreements (PTAs)"
author: "Cynthia Hester"
description: "DACSS 697E Social Networks Analysis Final Project"
date: "05/20/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories: 
  - Cynthia Hester
  - Final_Project
  - Preferential Trade Agreements (PTAs)  
  - Network of International Trade Agreements
---

```{r,include=FALSE}

#| label: setup
#| include: false

library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```


```{r,include=FALSE}

### Load necessary libraries

library(RColorBrewer)
library(tidyverse)
library(igraph)
library(intergraph)
library(tidyr)
library(statnet)
library(readr)
library(readxl)
library(kableExtra)
library(sna)
library(knitr)
library(network)
library(quarto)

```

# Research Question:

How does the network structure of Preferential Trade Agreements (PTAs) influence international trade patterns and economic integration among participating countries?



## Hypothesis:

H1: A more dense network structure in Preferential Trade Agreements (PTAs) leads to increased trade flow and economic integration among participating countries.


## Introduction:

In our Social Network Analysis course, Professor Rolfe highlighted the breadth of applications and emphasized the flexibility to examine any scenario that involved interactive systems. Taking this into account, I chose to focus my project on an area that piqued my personal interest - the intersection of travel and global trade.[^1]


**What is the source of the data?**

The data utilized for this project are derived from the Design of Trade Agreements (DESTA) database https://www.designoftradeagreements.org/project-description/, a comprehensive resource documenting numerous aspects of Preferential Trade Agreements (PTAs) globally, which was found via the World Trade Organization(WTO). With data spanning PTAs ratified since 1945, the DESTA database offers an examination of the various provisions and areas encompassed by these agreements. Utilizing this dataset was key in bringing the theories and methodologies of Social Network Analysis into practice in a real-world context.

**What is being measured**

The **Preferential Trade Agreement** (PTA) data measures the bilateral agreements between countries that provide preferential access to certain products from the countries involved, reducing tariffs or duties below the level that each country charges other members of the World Trade Organization (WTO).

The data includes variables, such as the (nodes) countries/regions involved, the type of agreement (free trade agreement, customs union, etc.), the date the agreement came into effect, and trade parameters.
*************************************************************************************

**Data Format**

It is provided as a matrix, where countries are listed both in rows and columns, and each cell in the matrix indicates whether a trade agreement exists between the two countries. Alternatively, it could be provided as an edgelist, where each row represents an agreement with two columns for the two countries involved. Sometimes, it might also be in the form of an affiliation network, representing memberships of different countries in different agreements. For the purposes of this project it is presented as a matrix.


## Data

As previously stated the data for this project was derived from the Design of Trade Agreements database (DESTA). The data utilized are a universe of cases.

## Wrangling and Exploration

#### Import DESTA data


```{r,results='hide'}

#preferential trade agreement data (pta)
#DESTA List of Treaties

excel_file <- "C:/Users/Bud/Downloads/desta_list_of_treaties_02_01_dyads.xlsx"

desta_list <- read_excel(excel_file, sheet = "data")

pta_data <- read_excel(path = excel_file, sheet = 2) 

pta_data$regioncon <- as.factor(pta_data$regioncon)


```
*************************************************************************************
## Inspect PTA trade agreement data/ Descriptive Statistics

```{r,results='hide'}
summary(pta_data)
```


```{r,results='hide'}

#first few rows of the data
head(pta_data) #look at the structure of the data

```


```{r,results='hide'}
#inspect the data
str(pta_data) 

```


```{r,results='hide'}
missing_values <- sapply(pta_data, function(x) sum(is.na(x)))
missing_values
```

```{r,results='hide'}
# Remove rows with missing values in country1 or country2 columns
pta_data <- pta_data[complete.cases(pta_data$country1, pta_data$country2), ]

```

```{r,results='hide'}
# Check for missing values in country1 and country2 columns
missing_country1 <- sum(is.na(pta_data$country1))
missing_country2 <- sum(is.na(pta_data$country2))

# Print the number of missing values
cat("Number of missing values in country1 column:", missing_country1, "\n")
cat("Number of missing values in country2 column:", missing_country2, "\n")

```
***********************************************************************************
#### Adjacency Matrix for entire dataset

```{r,results='hide'}
dim(pta_data)
```


```{r,results='hide'}

  pta_data[1 : 16, 1 : 16]
```


```{r,results='hide'}

#install and load the igraph package
if (!require(igraph)) {
    install.packages("igraph")
    library(igraph)
}


graph <- graph_from_data_frame(d = pta_data, directed = FALSE)

# extract the adjacency matrix
adj_matrix <- as_adjacency_matrix(graph)

# Print the adjacency matrix
print(adj_matrix)

```


nodes(countries)

```{r,results='hide'}
vcount(graph)
```

edges(trade ties)

```{r,results='hide'}
ecount(graph)
```

Calculating proportion of nodes 

```{r,results='hide'}
# Get the number of nodes
nodes <- V(graph)

# Calculate the proportion of non-zero entries
prop_nonzero <- sum(adj_matrix[] > 0) / length(nodes) ^ 2

# Print the proportion
print(prop_nonzero)
```
*******************************************************************************
**Analysis**

The output value 0.3540465 indicates that, based on the **pta_data**, approximately 35.4% of all potential bilateral Preferential Trade Agreements (PTAs) among countries are in place. This shows a considerable degree of interconnectedness in global trade. It also suggests substantial room for more PTAs, as 64.6% of potential agreements are unexplored.


```{r,results='hide'}


nodes <- data.frame(name = V(graph)$name,
                    all.degree = igraph::degree(graph), 
                    out.degree = igraph::degree(graph, mode = 'out'),
                    in.degree = igraph::degree(graph, mode = 'in'), 
                    total.degree = igraph::degree(graph, mode = 'total'),
                    eigen.centrality = igraph::eigen_centrality(graph)$vector)

head(nodes)


```
**********************************************************************************
**Distribution**

```{r,results='hide'}

library(ggplot2)

# histogram for all_degree
ggplot(nodes, aes(x=all.degree)) +
  geom_histogram(binwidth=1, fill="steelblue", color="black") +
  ggtitle("Distribution of All Degree") +
  xlab("All Degree") +
  ylab("Count")

# histogram for out_degree
ggplot(nodes, aes(x=out.degree)) +
  geom_histogram(binwidth=1, fill="steelblue", color="black") +
  ggtitle("Distribution of Out Degree") +
  xlab("Out Degree") +
  ylab("Count")

# histogram for in_degree
ggplot(nodes, aes(x=in.degree)) +
  geom_histogram(binwidth=1, fill="steelblue", color="black") +
  ggtitle("Distribution of In Degree") +
  xlab("In Degree") +
  ylab("Count")

# histogram for eigen_centrality
ggplot(nodes, aes(x=eigen.centrality)) +
  geom_histogram(binwidth=0.1, fill="steelblue", color="black") +
  ggtitle("Distribution of Eigen Centrality") +
  xlab("Eigen Centrality") +
  ylab("Count")

```


**Analysis**

This distribution reveals how nations in the PTA network interact. High degree centrality indicates a country's extensive trade relationships. High out-degree signifies proactive trade agreement initiation, while high in-degree suggests a country's attractiveness as a trade partner. Lastly, elevated eigenvector centrality reveals a nation's importance within major trade blocs or networks.


*************************************************************************************

## Dyad and Triad Census

```{r,results='hide'}

dyad.census(nodes)

```


```{r,results='hide'}
triad.census(nodes)
```



## Data Exploration

This is a large dataset so I will focus on two variables; **year and region**. To better understand the data I will visualize it using the ggplot package from the tidyverse.

```{r,results='hide'}

# counts the frequency of ties between any two countries formed by year and region.

pta_data_count <- pta_data %>% group_by(year, regioncon) %>% count(name) 
pta_data_count <- pta_data_count %>% arrange(desc(n))
kable(head(pta_data_count), col.names = c("Year", "Region", "Agreement name", "Number of Ties"),
      caption = "Frequency of Ties Table")


```

*******************************************************************************
## Centrality


Intercontinental Region

```{r,results='hide'}

pta_intercon <- pta_data %>% filter(regioncon == "Intercontinental")
pta_intercon_net <- graph_from_edgelist(cbind(pta_intercon$country1, pta_intercon$country2), directed = F)
```





 **Degree Centrality Intercontinental Region**

```{r}
library(igraph)



# Create a graph from the pta_data
pta_net <- graph_from_data_frame(d = pta_data_count, directed = TRUE)

# Calculate the degree centrality
degree_centrality <- igraph::degree(pta_net, mode = "all")

# Print the degree centrality values
print(degree_centrality)




```

**Degree Centrality Analysis**

In the context of degree centrality, the nodes with the highest degree centrality values can be considered the most important or influential nodes in the network. These nodes have a large number of connections (edges) with other nodes in the network. 

Nodes(regions) with the highest degree centrality values are:

- Europe: 260
- Americas: 231
- Asia: 140
- Intercontinental: 382

These nodes represent regions or continents and have the highest number of connections with other nodes in the network. Therefore, they can be considered the most significant or influential  network connectivity.


**Betweenness Centrality**

```{r}

library(igraph)

# graph from the pta_data_count
pta_net <- graph_from_data_frame(d = pta_data_count, directed = F)

# betweenness centrality
betweenness_centrality <- igraph::betweenness(pta_net, directed = F)

# Print the betweenness centrality values
print(betweenness_centrality)

```



**Betweenness Analysis**

Egypt, Jordan, and Morocco are identified as nodes with high betweenness centrality. These countries serve as key intermediaries in facilitating trade relationships and information exchange between other countries within the network.

On the other hand, nodes with lower betweenness centrality values, such as Albania or Armenia, have less influence in connecting different parts of the network. While they still contribute to the network structure, their impact on information flow and communication may be comparatively lower.


*******************************************************************************


**Summary of Preferential Trade Agreement data**

```{r,results='hide'}
kable(summary(pta_data_count), col.names = c("Year", "Region", "Agreement name", "Number of Ties"), 
      caption = "A Summary of the Frequency of Ties Data") #summary() provide a summary of the data
```

##  Overview

An overview of the data reveals six distinct categories: Africa, Americas, Asia, Europe, Oceania, and Intercontinental. Interestingly, most PTAs are Intercontinental, tallying 382 agreements, while Oceania, given the limited number of nations in the region, has only  12 intra-regional agreements.

The next stage of my exploration will involve identifying patterns of multilateral and bilateral PTA formation among different countries.


```{r,results='hide'}

#filter the new data: which countries tend to form multilateral PTAs and which countries tend to form bilateral PTAs

pta_data_count <- pta_data_count %>% mutate(bilat = ifelse(n == 1, "bilateral", "multilateral")) #creating a new variable to see if a tie belongs to a bilateral or multilateral treaty

#png(filename = "network1.png") # save as png
pta_data_count$bilat <- as.factor(pta_data_count$bilat)
region_count <- pta_data_count %>% group_by(regioncon, bilat) %>% count(regioncon)
ggplot(region_count, aes(x= regioncon, y = n, color = bilat, fill = bilat)) +
  geom_col()

```


It's observed that African nations exhibit a preference for multilateral **PTAs** as opposed to bilateral ones. Conversely, countries in the Americas and Asia predominantly lean towards bilateral **PTAs**. Europe and Oceania display an almost equal inclination towards both bilateral and multilateral **PTAs**. Of note when it comes to inter-regional PTAs, there appears to be no clear preference between bilateral and multilateral arrangements.


*******************************************************************************


Creating a plot to visualize the number of ties formed annually by region reinforces indicates thats Africa is particularly prominent in terms of the number of ties established, as well as in forming intra-regional PTAs, with Europe following closely. Meanwhile, Asia, the Americas, and Oceania demonstrate a preference for forming bilateral agreements.


```{r,results='hide'}
#Visualizing ties formed annually by global region
#png(filename = "network2.png") # save as png
ggplot(pta_data_count, aes(x = year, y = n, color = regioncon, fill = regioncon)) +
  geom_col() + facet_wrap(~regioncon)

```

Indegree/ Outdegree of PTAs Intercontinental Region

```{r}
# indegree and outdegree
indegree <- colSums(as.matrix(as_adjacency_matrix(pta_intercon_net)))
outdegree <- rowSums(as.matrix(as_adjacency_matrix(pta_intercon_net)))

# data frame with indegree and outdegree values
degree_df <- data.frame(country = V(pta_intercon_net)$name, indegree = indegree, outdegree = outdegree)

# Print the degree information
print(degree_df)

```

**Analysis**

We see that countries like Egypt, Jordan, Morocco, and Tunisia have relatively high indegree values, indicating that they receive a substantial number of connections from other countries.

On the other hand, outdegree represents the number of outgoing connections or links originating from a particular country. Higher outdegree values suggest that the country has more connections with other countries in the network. Countries like Belgium, Denmark, France, and Turkey have notable outdegree values, indicating that they have numerous connections with other countries.



*******************************************************************************


## Creating a Network of Intercontinental Regions

```{r,results='hide'}
# rechecking the first 6 rows of the data
head(pta_data)

```


We create a network using an edgelist as the data structure, allowing us to visualize the connections between global regions and the formation of PTAs. This approach enables us to analyze the structural patterns within the global PTA landscape.

```{r,results='hide'}
temp_graph <- graph_from_edgelist(cbind(head(pta_data$country1), head(pta_data$country2)), directed = F)
get.adjacency(temp_graph)
```


**Edgelist**

```{r,results='hide'}
cbind(head(pta_data$country1), head(pta_data$country2))
```

```{r,results='hide'}
# using the igraph package to create an edgelist

edge_net <- graph_from_edgelist(cbind(head(pta_data$country1), head(pta_data$country2)),
directed = FALSE) 

#png(filename = "network3.png") # save as png

set.seed(123) #randomizes the configuration of the plot


plot.igraph(edge_net)
```


**Subset of Adjacency matrix**

The adjacency matrix of the Preferential Trade Agreement (PTA) dataset provides a  representation of the bilateral trade relationships between countries participating in trade agreements. Each entry in the matrix indicates whether a specific pair of countries has established a preferential trade agreement or not. This binary representation allows us to identify the specific countries that are directly connected through these agreements.

By examining the adjacency matrix, we can observe the clustering and groupings of countries that have formed preferential trade agreements. These clusters indicate the presence of regional trade blocs or alliances, where countries within each cluster have established closer trade ties among themselves. Conversely, countries that are not connected in the adjacency matrix may have limited or no preferential trade agreements with other countries in the dataset.


```{r,results='hide'}
get.adjacency(edge_net)
```


```{r,results='hide'}

matrix_net <- graph_from_adjacency_matrix(get.adjacency(edge_net), mode = "undirected")
set.seed(123) 
plot.igraph(matrix_net)


```

Here we see that both graphs appear identical. This is attributed to the edge list and adjacency matrix representing an undirected network. In an undirected network, the connections between nodes are symmetric and bidirectional.

Network representation of countries based on their inter-regional Preferential Trade Agreements (PTAs)

*********************************************************************************





This code enhances the Preferential Trade Agreement(PTA) network graph by attributing each trade agreement with its respective name and the year it was established.


```{r,results='hide'}
#The edges of the network are referred to as E$agt_name, representing the names of the PTAs, and E$year, representing the respective years of these PTAs.

E(pta_intercon_net)$agt_name <- pta_intercon$name
E(pta_intercon_net)$year <- pta_intercon$year

```

**Analysis**

Of note, there may be duplicate edges between pairs of countries, indicating multiple PTAs formed between them. To address this, I assign a weight to each pair based on the number of shared PTAs. To simplify the network by removing loops and consolidating duplicate edges, I utilize the simplify() function from the igraph package. This process entails aggregating the weights of duplicate edges, ensuring a streamlined representation of the network.

**Weights**

```{r,results='hide'}

# Assigning weight 1 to all edges
E(pta_intercon_net)$weight <- 1


pta_intercon_simp <- graph.strength(pta_intercon_net, mode = "all", loops = FALSE)


```


**Weights Discussion**

I first normalize the relationships within the Preferential Trade Agreement (PTA) interconnection network. I do this by assigning a uniform weight of 1 to all edges, effectively treating all connections as equal. Following this, I calculate the strength, or degree, of each node in the network, which gives me the number of connections each node, or country, has within this network. In this calculation, I treat the network as undirected and exclude any loops. The resulting data offers a straightforward view of the interconnectedness of each country within the PTA network, under the assumption that all trade agreements hold the same importance or weight.




#### Vertices and Edges

```{r,results='hide'}

head(V(pta_intercon_net)$name)

```

**Vertices and Edges**

number of vertices,and edges in the network

```{r,results='hide'}
summary(pta_intercon_net)
```


```{r}
print(vcount(pta_intercon_net))
```



The summary indicates the number of **vertices(nodes or countries)** **(191)** and edges **(9316)** in the graph. The following line presents a list of attributes categorized by their kind (g for graph, v for vertex, and e for edge).

********************************************************************************

## Network Visualizations

We will generate two separate plots: one depicting the Inter-regional PTA network consisting exclusively of Inter-regional PTAs, and another illustrating the Bilateral PTA network that includes Bilateral agreements regardless of whether they are Inter-regional or regional in nature.



```{r,results='hide'}

pta_bilat <- pta_data %>% filter(typememb == 1)
pta_bilat_net <- graph_from_data_frame(d = pta_bilat, directed = FALSE)

E(pta_bilat_net)$agt_name <- pta_bilat$name
E(pta_bilat_net)$year <- pta_bilat$year

E(pta_bilat_net)$weight <- 1
pta_bilat_simp <- simplify(pta_bilat_net)


```


Network depiction of Inter-regional Preferential Trade Agreements (PTA)

```{r,results='hide'}
set.seed(123)
plot.igraph(pta_intercon_net,
            vertex.size = 5,
            vertex.color = "red",
            vertex.label = V(pta_intercon_net)$name,
            vertex.label.cex = .5,
            vertex.frame.color = NA,
            layout = layout.fruchterman.reingold,
            main = "Network of Inter-regional Preferential Trade Agreements(PTA)"
            )
```
Using all of the data depicts a "hairball" I will therefore use a subset of the data for more comprehensible data representation.

```{r,results='hide'}

E(pta_bilat_simp)$color <- ifelse(E(pta_bilat_simp)$weight > 1,'blue', 'grey')
plot.igraph(pta_bilat_simp,
            vertex.size = 5,
            vertex.color = "red",
            vekrtex.label = V(pta_bilat_simp)$name,
            vertex.label.cex = .5,
            vertex.frame.color = NA,
            layout = layout.fruchterman.reingold,
            edge.width = E(pta_bilat_simp)$weight,
            main = "Network of Bilateral Preferential Trade Agreements(PTA)"
            )




```


```{r,results='hide'}

# Bar plot of Bilateral network

E(pta_intercon_net)$bilateral <- ifelse(pta_intercon$typememb == 1, 1,0)
bilateral_var <- as.data.frame(E(pta_intercon_net)$bilateral)
ggplot(bilateral_var, aes(x =E(pta_intercon_net)$bilateral)) +
  geom_bar(fill = "blue", color = "green") +
  theme_minimal() +
  labs(x = "Bilateral/Multilateral Inter-Regional Preferential Trade Agreements (PTA)", y = "Count")

#png(filename = "network4.png") # save as png

```

The bar plot shows the distribution and relative importance of bilateral and multilateral Preferential Trade Agreements (PTAs) within the framework of social network analysis applied to inter-regional trade relationships. Notably, the left bar exhibits a considerably higher frequency of PTAs compared to the right bar, indicating a distinct prevalence of bilateral PTAs in the network.
Conversely,the lower frequency of multilateral PTAs on the right bar signifies a relatively reduced presence of agreements involving multiple nations in the network. Multilateral PTAs typically require more extensive negotiations and coordination among a larger number of countries. The scarcity of such agreements suggests a less pronounced emphasis on broader inter-regional collaboration within the analyzed trade network.

******************************************************************************
```{r,results='hide'}

#bilateral inter-regional PTAs
pta_intercon_bilat <- subgraph.edges(pta_intercon_net,which(E(pta_intercon_net)$bilateral == 1))
plot(pta_intercon_bilat,
     vertex.size = 5,
     vertex.color = "red",
     vertex.label = V(pta_intercon_bilat)$name,
     vertex.label.cex = .5,
     vertex.frame.color = NA,
     layout = layout.fruchterman.reingold,
     main = "Network of Bilateral Inter-regional PTAs")

#(filename = "network5.png") # save as png
```




**Eigen Centrality Intercontinental Region**

I plot an Eigen centrality for the intercontinetal region which quantifies a country's influence by factoring both the quantity and quality of its trade agreements.

```{r,results='hide'}

# eigen centrality 
V(pta_intercon_bilat)$eigen_cen <- eigen_centrality(pta_intercon_bilat)$vector

print((V(pta_intercon_bilat)))


eigen_cen <- eigen_centrality(pta_intercon_bilat)$vector

# bar plot

barplot(eigen_cen, names.arg = V(pta_intercon_bilat)$name, xlab = "Country", ylab = "Eigen Centrality", col = "blue")

title(main = "Eigen Centrality of Countries in PTA Intercontinental Bilateral Network")
#png(filename = "network8.png") # save as png
```

**Analysis**

This bar plot represents the eigen centrality values of countries in the PTA Intercontinental Bilateral network. This visualization helps to analyze and interpret the relative importance of countries based on their connections within the network. Countries with higher eigen centrality scores are considered more influential and have stronger trade relationships, while those with lower scores have less impact. This helps identify key players and the distribution of power within the PTA network.



********************************************************************************

```{r,results='hide'}


E(pta_intercon_bilat)$weight <- 1
pta_intercon_bilat_sim <- pta_intercon_bilat %>% simplify(edge.attr.comb = list(weight = "sum"))
plot(pta_intercon_bilat_sim,
     vertex.size = 5,
     vertex.color = "blue",
     vertex.label = V(pta_intercon_bilat_sim)$name,
     vertex.label.cex = .5,
     vertex.frame.color = NA,
     edge.width = E(pta_intercon_bilat_sim)$weight,
     layout = layout.fruchterman.reingold,
     main = "Network of Bilateral Inter-regional Preferential Trade Agreements(PTAs)
     ")

#png(filename = "network6.png") # save as png
```


*******************************************************************************

## Density and Transitivity

```{r,results='hide'}
bilat_density <- edge_density(pta_intercon_bilat)
bilat_transitivity <- transitivity(pta_intercon_bilat)

#density
pta_density <- edge_density(pta_intercon_net)
pta_transitivity <- transitivity(pta_intercon_net)

bilat_density

```



```{r,results='hide'}


pta_bilat_2 <-  subgraph.edges(pta_bilat_simp,which(E(pta_bilat_simp)$weight > 1))
plot.igraph(pta_bilat_2,
            vertex.size = 5,
            vertex.color = "green",
            vekrtex.label = V(pta_bilat_2)$name,
            vertex.label.cex = .5,
            vertex.frame.color = NA,
            layout = layout.fruchterman.reingold,
            edge.width = E(pta_bilat_2)$weight,
            main = "Countries with more than One Bilateral PTAs"
            )
png(filename = "network11.png") # save as png

```

********************************************************************************
## Community

**Walktrap Algorithim**

For this project examining the Preferential Trade Agreement (PTA) network, I use the Walktrap algorithm, we examined in the 7th tutorial. This approach seeks to find communities within this complex trade network by employing random walks. The guiding principle behind this method is that shorter, random walks will likely stay within the confines of a single community. By applying the Walktrap algorithm to the PTA network, I aim to discern the community structures present within this global trade landscape, thereby shedding light on the intricate dynamics and patterns that shape international trade agreements.


```{r,results='hide'}

intercon_bilat_walktrap <- cluster_walktrap(pta_intercon_bilat_sim)
membership(intercon_bilat_walktrap)
```


```{r,results='hide'}
#bilateral walktrap
sizes(intercon_bilat_walktrap)

```




```{r,results='hide'}


# walktrap algorithim output data is in a vector named "output"
output <- c(11, 14, 18, 4, 2, 20, 3, 12, 2)

# data frame
output_df <- data.frame(
  "Community" = paste0("Community_", 1:length(output)),
  "Size" = output
)

# Print the data frame
print(output_df)

```


**Preferential Trade Agreement(PTA) Walktrap Analysis**

The Walktrap algorithm's output portrays the distribution of nodes across various communities in the Preferential Trade Agreement (PTA) network. Each community number is paired with a value that signifies the count of nodes (or countries) within that specific community.

Here is what we can infer:

- Community 1 consists of 11 nodes.
- Community 2 has 14 nodes.
- Community 3 comprises 18 nodes.
- Community 4 includes 4 nodes.
- Community 5 houses 2 nodes.
- Community 6 holds 20 nodes.
- Community 7 encapsulates 3 nodes.
- Community 8 encompasses 12 nodes.
- Community 9 shelters 2 nodes.

This distribution provides a snapshot of the diverse interconnectedness among global trade relationships in the PTA network. Communities with larger node counts might represent robustly interconnected trading partnerships, while smaller communities may indicate more specific or self-contained trade relationships.

## Density 

**PTA Bilateral Density Analysis**

The value 0.04924761 represents the bilateral density in the context of the Preferential Trade Agreement (PTA) dataset. Network density measures the actual connections in a network relative to the total possible connections. 

Here, a density of 0.04924761 suggests that nearly 4.92% of all potential bilateral trade agreements among countries are realized. This means, on an average, a country in the network has a trade agreement with approximately 4.92% of all other potential trading partners. Given that this value is relatively low, it indicates that the bilateral trade network is sparsely connected, suggesting the potential for many more trade agreements to be formed.

## Transitivity

**PTA Bilateral Transitivity**

```{r,results='hide'}
bilat_transitivity
```
**Analysis**

In the context of the Preferential Trade Agreement (PTA) dataset, a bilateral transitivity of 0.1098987 signifies that about 11% of the potential triangular trade relationships among countries are realized. This implies that if two countries both have a PTA with a third one, there is an approximately 11% probability that these two countries will also have a trade agreement directly with each other.

This relatively low transitivity index suggests that the formation of PTAs is not extensively based on the principle of transitivity. Rather, it underscores that the establishment of PTAs among countries may be influenced by a multitude of factors, including geographical proximity and political or economic dynamics.


## Dendrogram

**Dendrogram**

In the context of the Preferential Trade Agreement (PTA) dataset focusing on the variables, the dendrogram emerged as an influential graphical representation. It conveys the nested arrangement of relationships among countries or corresponding trade agreements, an arrangement that is guided by specified attributes or significant performance indicators. This hierarchical classification reveals nuanced patterns in international trade interactions, which contributes to a better understanding of intricate PTAs. 

```{r,results='hide'}

dist_matrix <- dist(pta_data) 

# Hierarchical clustering using complete linkage
hc <- hclust(dist_matrix, method = "complete")

# Plot dendrogram
plot(hc, hang = -1, cex = 0.6, main = "Dendrogram for PTA Data")
#png(filename = "network9.png") # save as png
```


********************************************************************************


## Discussion

The Preferential Trade Agreement (PTA) dataset, spanning from 1958 to 2021, offers a comprehensive view into the evolution of global trade relationships. Delving into this dataset  has afforded me a richer understanding of social network analysis in the context of international trade. This journey, albeit marked by a steep learning curve, has been an enlightening one.

The dominance of bilateral PTAs suggests that countries in the network tend to engage in preferential trade arrangements on a bilateral basis, forming direct connections with specific trading partners. This finding underscores the significance of bilateral relationships in shaping the overall structure of the inter-regional trade network. Bilateral PTAs are often favored due to their simplicity in negotiation and implementation, enabling countries to establish tailored agreements that cater to their specific needs and preferences.[^4] These agreements facilitate focused interactions between pairs of countries, fostering closer economic ties and potentially enhancing trade efficiency and cooperation.

*************************************************************************************


## Conclusion/Reflection


Prior to this project, the critical role that variable selection plays in shaping network interpretation was an aspect I hadn't fully appreciated. Now, I understand that even with the PTA dataset, or for any social network analysis, significant effort can be poured into understanding the relationships between just a couple of variables to derive meaningful insights. Upon reflection, the prospect of analyzing data from more recent years to assess how key features like **node centrality,transitivity,density** and community structure have evolved in the midst of global challenges like the COVID-19 pandemic would merit future study, especially considering recent challenges in global trade and supply chain restrictions.

Despite the hurdles encountered, the process of dissecting this dataset has been rewarding. Every decision made echoed in the analysis outcomes, underscoring the potential for subtle shifts to lead to major interpretation changes. I plan to continue my exploration of the PTA data and branch out to related projects to enhance my understanding of Social Network Analysis. 


**********************************************************************************

## References/Resources

[^1]: https://www.designoftradeagreements.org/project-description/
[^2]: Wasserman, S., & Faust, K. (1994). Social network analysis: Methods and 
      applications. Cambridge University Press.
[^3]: Newman, M. E. J. (2018). Networks. Oxford University Press.     
[^4]: https://www.wto.org/
[^5]: Robins, G., Pattison, P., Kalish, Y., & Lusher, D. (2007). An Introduction to Exponential Random Graph (p*) Models for Social Networks. Social Networks, 29, 172-191. 










