---
title: "Week1_Challenge"
author: "Rahul Gundeti"
description: "Loading Data and Creating a Network"
date: "02/13/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
# editor: visual
categories: 
  - challenge_1
  - instructions
---

```{r}
#| label: setup
#| include: false
```

#### Challenge Overview

This Challenge is to

1\) Read a dataset, and

2\) Create a network object

#### Loading the data

Read in one (or more) of the following data sets, using the correct R package and command.

\- got_marriages.csv

\- fish_encounters dataset (available in the \`tidyr\` package)

\- got_like_dislike.csv

```{r}
library(readr)
got_like_dislike <- read_csv("_data/got/got_like_dislike.csv")
head(got_like_dislike)
```

#### Creating a network

Load the package \`igraph\` and create an \`igraph\` object (i.e. a graph or network) in the form of an edge list. The command may vary whether the data is given as a list of connections or an adjacency matrix. Is the network directed or undirected; weighted or unweighted; unimodal or bipartite? Can you plot it?

```{r}
library(igraph)
net.igraph <- graph_from_data_frame(got_like_dislike, directed = TRUE)

# Check if the network is directed or undirected, weighted or unweighted, unimodal or bipartite
is_bipartite(net.igraph)
is_weighted(net.igraph) 
is_directed(net.igraph) 
plot(net.igraph)
```
